\chapter*{Preface}
\setheader{Preface}

The big hype around Bitcoin and other cryptocurrencies seems to create a distorted picture of the 
original idea of a distributed, democratic financial system. The general public does not appreciate
the simplicity and brilliance of the concept but rather the price swings and the
possibility to become rich. 

A similar distortion happened to the internet itself. Once envisioned as an 
open and free network for information exchange, the World Wide Web has been transformed into an
infrastructure for commerce, data collection and cyber attacks. Certainly, many companies have created 
solutions that simplify our lives significantly. But the free use of all kinds of services comes at 
an invisible cost. Slowly, that cost becomes more visible as targeted misinformation campaigns are
dismantling democracies and data collection policies of the big internet monopolies are uncovered.

Decentralization is a natural solution to this problem. Some fundamental services such as transferring
money, communication and the access to information should not be controlled by single entities and 
exploited for profit. In the same sense, creating trust with other internet users should be 
enabled through a decentralized system. This is the motivation to contribute to the design of a
distributed trust system.

Working on this thesis has been a journey through many fields of science, from evolution theory and 
social sciences to synchronization methods of replicated databases. 
It has in many ways opened my eyes to the problems of centralized 
organizations and the advantages that distributed systems offer. Also, it sparked my interest in 
trust, cooperation and their relation to problems of society.

I would like to thank my supervisor Dr. Johan Pouwelse for his enthusiasm, infinitely large pool of
ideas and suggestions as well as his invaluable feedback. I thank the Tribler team for keeping me 
awake with coffee and the review work of Quinten and Martijn. I thank my parents, my girlfriend, Lyla, 
and my closest friends for their support in moments of self-doubt. A special thanks goes to my 
housemates Marnix and Liesette who helped me through the last weeks. I would also like to thank 
Alexander for our interesting talks at lunch. Without you it wouldn't have been possible.


\begin{flushright}
{\makeatletter\itshape
    \@author \\
    Delft, August 2018
\makeatother}
\end{flushright}

